### 2024-03-29 scripts/drop.long.line.Rmd
made boat type data plots

---
title: "drop.long.line"
author: "daniellembecker"
date: "2024-03-29"
output: html_document
---

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}

# Load dataframe
boat_data <- read.csv("data/dropline.longline.csv")

# Define the order of size categories
boat_data$size_cat <- factor(boat_data$size.cat, levels = c('Nano', 'Small', 'Medium', 'Large'))

# make stacked bar plot
stacked <- ggplot(boat_data, aes(x = size_cat)) +
  geom_bar(aes(y = dropline), fill = 'blue', width = 0.35, stat = "identity") +
  geom_bar(aes(y = longline), fill = 'lightblue', width = 0.35, stat = "identity") +
  labs(x = 'Size Category', y = 'Number of Boats', title = 'Number of Boats by Size Category') +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0));stacked   # Remove space between bars and x-axis

# Save the plot
ggsave("output/stacked_plot.png", plot = stacked, width = 8, height = 6, dpi = 300)

#make not stacked bar plot

# Reshape the data from wide to long format
library(reshape2)
boat_data_long <- melt(boat_data, id.vars = "size_cat", measure.vars = c("dropline", "longline"))
boat_data_long$variable <- factor(boat_data_long$variable, levels = c("dropline", "longline"))

# Plot the data
bar.plot <- ggplot(boat_data_long, aes(x = size_cat, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.4), width = 0.35) +
  labs(x = 'Size Category', y = 'Number of Boats', title = 'Number of Boats by Size Category',
       fill = "Boat Type") +
  scale_fill_manual(values = c("blue", "lightblue"), labels = c("Dropline", "Longline")) +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)); bar.plot

# Save the plot
ggsave("output/unstacked.plot.png", plot = bar.plot, width = 8, height = 6, dpi = 300)

```