### 2024-04-12 scripts/data_collation.Rmd
finished cleanign data

---
title: "data_collation"
author: "daniellembecker"
date: "2024-03-21"
output: html_document
---

# Use this script to clean and merge unique data types (e.g., swath, fish, UPC) from multiple years

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(lubridate)
```

# Load all data files from multiple years and seperate them into specific categories: UPC, swath, Morph, Site, QUAD, kelp, fish, and COVER
```{r}

# Set your directory path
directory <- "data/kelp_timeseries_data/"

# List files in the directory
files <- list.files(directory)

# Initialize an empty list to store data frames
data_frames <- list()

# Loop through each file
for (file in files) {
    # Read the file into a data frame
    # Adjust the read function according to your file type (e.g., read.csv, read.table)
    # Specify other parameters as needed (e.g., header = TRUE)
    file_path <- file.path(directory, file)
    df_name <- gsub("\\..*", "", file)  # Extract file name without extension
    assign(df_name, read.csv(file_path, stringsAsFactors = FALSE))
}

# Now, each file is a separate data frame with its own variable name

```

# Combine kelp dataframes
```{r}
# Kelp data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)

#delete repetative data collector columns
kelp_data <- kelp_data[ -c(11:17) ]

# Use lubridate to format date to be consistent
kelp_data$DATE <- mdy(kelp_data$DATE)

# Rename King's Beach to Kings Beach
kelp_data$SITE[kelp_data$SITE == "King's Beach"] <- "Kings Beach"

# Rename ft. weatherill to fort wetherill
kelp_data$SITE[kelp_data$SITE == "Ft Weatherill"] <- "Fort Wetherill"

# output updated dataframe
write.csv(kelp_data, "output/cleaned_data/kelp.data.clean.csv", row.names = FALSE)
```

# Combine Kelp_Morph dataframes
```{r}

# Morph data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)

# remove columns with observer and unneeded data
morph_data <- morph_data[ -c(11:14) ]

# Make all columns capitalized
colnames(morph_data) <- toupper(colnames(morph_data))

# Use lubridate to format date to be consistent
morph_data$DATE <- mdy(morph_data$DATE)

# output updated dataframe
write.csv(morph_data, "output/cleaned_data/morph.data.clean.csv", row.names = FALSE)
```

# Combine fish dataframes
```{r}

# Fish data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes)

# remove columns with observer and unneeded data
fish_data <- fish_data[ -c(10:14) ]

# Use the actual size data to bin into the size classes to combine with other dataframe
fish$SIZE <- cut(fish$actual_size, breaks = c(0, 10, 20, 30, 40, 50, 70, 100, Inf),
               labels = c("< 10", "10_20", "20_30", "30_40", "40_50", "50_70", "70_100", "> 100"),include.lowest = TRUE, right = FALSE)

# combine updated dataframes
fish_data <- left_join(fish, fish_data)

# remove columns with VIS
fish_data <- fish_data[ -c(14) ]

# Use lubridate to format date to be consistent
fish_data$DATE <- mdy(fish_data$DATE)

# output updated dataframe
write.csv(fish_data, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)

# look at fish_Aggregated later - Austin, this is kelp density possibly?

```

# Combine UPC dataframes
```{r}

# UPC data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
UPC_data <- rbind.fill(UPC, UPC_2017, UPC_2018, FtWeatherill_UPC_2016, FtWeatherill_UPC_2016_CT)

# Rename ft. wetherill to fort wetherill
UPC_data$SITE[UPC_data$SITE == "Ft. Weatherill"] <- "Fort Wetherill"

# Rename fort weatherill
UPC_data$SITE[UPC_data$SITE == "Fort Weatherill"] <- "Fort Wetherill"

# remove columns with observer and unneeded data
UPC_data <- UPC_data[ -c(21:24) ]

# Make all columns capitalized
colnames(UPC_data) <- toupper(colnames(UPC_data))

# Use lubridate to format date to be consistent
UPC_data$DATE <- mdy(UPC_data$DATE)

# output updated dataframe
write.csv(UPC_data, "output/cleaned_data/UPC.data.clean.csv", row.names = FALSE)

#UPC data frame does not have month and day so that is why it isn't connecting fully
```

# Combine site.info dataframes
```{r}

# site information data is cleaned and ready

# Make all columns capitalized
colnames(Site_info_2018) <- toupper(colnames(Site_info_2018))

# combine site collection data
site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)

# remove columns with observer and unneeded data
site.info.dat <- site.info.dat[ -c(12:15) ]

# output updated dataframe
write.csv(site.info.dat, "output/cleaned_data/site.info.data.clean.csv", row.names = FALSE)

# Austin - the site info for 2018 has a start and stop latitude so don't know what to do here

```

# Combine swath dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
swath_data <- rbind.fill(swath, swath_2018, swath_count, FtWeatherill_swath_2016)

# Rename ft. wetherill to fort wetherill
swath_data$SITE[swath_data$SITE == "Ft.Weatherill"] <- "Fort Wetherill"

# swath dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes

# load species names and species code dataframe
species.metadata <- read.csv("data/species_code_metadata.csv")

# Convert species names to codes
for (col in names(swath)) {
  swath[, col] <- species.metadata$code[match(swath[, col], species.metadata$species.substrate)]
}

#some of the species in swath dataframe are common names and species metadata in species names

```

# Combine QUAD dataframes
```{r}

# To do: species codes transform

# use rbind.fill() method in R to combine data frames with different columns
QUAD_data <- rbind.fill(QUAD, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

# Rename ft. wetherill to fort wetherill
QUAD_data$SITE[QUAD_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"

```