### 2024-05-03 scripts/data_collation.Rmd
almost finished organizing swaath data

---
title: "data_collation"
author: "daniellembecker"
date: "2024-03-21"
output: html_document
---

# Use this script to clean and merge unique data types (e.g., swath, fish, UPC) from multiple years

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(lubridate)
```

# Load all data files from multiple years and seperate them into specific categories: UPC, swath, Morph, Site, QUAD, kelp, fish, and COVER
```{r}

# Set your directory path
directory <- "data/kelp_timeseries_data/"

# List files in the directory
files <- list.files(directory)

# Initialize an empty list to store data frames
data_frames <- list()

# Loop through each file
for (file in files) {
    # Read the file into a data frame
    # Adjust the read function according to your file type (e.g., read.csv, read.table)
    # Specify other parameters as needed (e.g., header = TRUE)
    file_path <- file.path(directory, file)
    df_name <- gsub("\\..*", "", file)  # Extract file name without extension
    assign(df_name, read.csv(file_path, stringsAsFactors = FALSE))
}

# Now, each file is a separate data frame with its own variable name

```

# Combine kelp dataframes
```{r}
# Kelp data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)

#delete repetative data collector columns
kelp_data <- kelp_data[ -c(11:17) ]

# Use lubridate to format date to be consistent
kelp_data$DATE <- mdy(kelp_data$DATE)

# Rename King's Beach to Kings Beach
kelp_data$SITE[kelp_data$SITE == "King's Beach"] <- "Kings Beach"

# Rename ft. weatherill to fort wetherill
kelp_data$SITE[kelp_data$SITE == "Ft Weatherill"] <- "Fort Wetherill"

# output updated dataframe
write.csv(kelp_data, "output/cleaned_data/kelp.data.clean.csv", row.names = FALSE)
```

# Combine Kelp_Morph dataframes
```{r}

# Morph data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)

# remove columns with observer and unneeded data
morph_data <- morph_data[ -c(11:14) ]

# Make all columns capitalized
colnames(morph_data) <- toupper(colnames(morph_data))

# Use lubridate to format date to be consistent
morph_data$DATE <- mdy(morph_data$DATE)

# output updated dataframe
write.csv(morph_data, "output/cleaned_data/morph.data.clean.csv", row.names = FALSE)
```

# Combine fish dataframes
```{r}

# Fish data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes)

# remove columns with observer and unneeded data
fish_data <- fish_data[ -c(10:14) ]

# Use the actual size data to bin into the size classes to combine with other dataframe
fish$SIZE <- cut(fish$actual_size, breaks = c(0, 10, 20, 30, 40, 50, 70, 100, Inf),
               labels = c("< 10", "10_20", "20_30", "30_40", "40_50", "50_70", "70_100", "> 100"),include.lowest = TRUE, right = FALSE)

# combine updated dataframes
fish_data <- left_join(fish, fish_data)

# remove columns with VIS
fish_data <- fish_data[ -c(14) ]

# Use lubridate to format date to be consistent
fish_data$DATE <- mdy(fish_data$DATE)

# output updated dataframe
write.csv(fish_data, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)

# look at fish_Aggregated later - Austin, this is kelp density possibly?

```

# Combine UPC dataframes
```{r}

# UPC data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
UPC_data <- rbind.fill(UPC, UPC_2017, UPC_2018, FtWeatherill_UPC_2016, FtWeatherill_UPC_2016_CT)

# Rename ft. wetherill to fort wetherill
UPC_data$SITE[UPC_data$SITE == "Ft. Weatherill"] <- "Fort Wetherill"

# Rename fort weatherill
UPC_data$SITE[UPC_data$SITE == "Fort Weatherill"] <- "Fort Wetherill"

# remove columns with observer and unneeded data
UPC_data <- UPC_data[ -c(21:24) ]

# Make all columns capitalized
colnames(UPC_data) <- toupper(colnames(UPC_data))

# Use lubridate to format date to be consistent
UPC_data$DATE <- mdy(UPC_data$DATE)

# output updated dataframe
write.csv(UPC_data, "output/cleaned_data/UPC.data.clean.csv", row.names = FALSE)

#UPC data frame does not have month and day so that is why it isn't connecting fully
```

# Combine site.info dataframes
```{r}

# site information data is cleaned and ready

# Make all columns capitalized
colnames(Site_info_2018) <- toupper(colnames(Site_info_2018))

# combine site collection data
site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)

# remove columns with observer and unneeded data
site.info.dat <- site.info.dat[ -c(12:15) ]

# output updated dataframe
write.csv(site.info.dat, "output/cleaned_data/site.info.data.clean.csv", row.names = FALSE)

# Austin - the site info for 2018 has a start and stop latitude so don't know what to do here

```

# Combine swath dataframes
```{r}

# swath dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes

# Load species names and species code data frame
species.metadata <- read.csv("data/species_code_metadata.csv")

# Function to rename columns using metadata
# Get the unique column names from the swath dataframe
column_names <- colnames(swath)

# Create a lookup table with name-to-code mappings
name_to_code <- setNames(species.metadata$code, species.metadata$name)

# Rename the columns using the lookup table
  new_column_names <- sapply(column_names, function(name) {
    if (name %in% names(name_to_code)) {
      return(name_to_code[name])
    } else {
      return(name)
    }
  })

# Rename the columns
colnames(swath) <- new_column_names

# Convert data in numeric columns to numeric
swath.con <- swath %>%
  mutate(across(AGH:BLOA, as.numeric))

# Melt only columns AGH:WHEL into long format
swath_long <- swath.con %>%
  pivot_longer(cols = AGH:BLOA, names_to = "SP_CODE", values_to = "value")

# Need to now move values to binned swatch categories as columns
# Pivot the dataframe into wide format
swath_binned <- swath_long %>%
  pivot_wider(names_from = swath, values_from = value)

# make columns lowercase
# Get the current column names
current_names <- colnames(swath_binned)

# Convert the column names to lowercase while preserving the first letter
new_names <- sapply(current_names[1:6], function(x) {
  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
})

# Assign the new column names to the first six columns of the dataframe
colnames(swath_binned)[1:6] <- new_names

# rename swatch category columns in swath count and swath 208 dataframe
swath_2018 <- swath_2018 %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

swath_count <- swath_count %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# Remove four columns in a specific order swath_2018
swath_2018 <- swath_2018 %>%
  select(-12:-15)

# Remove four columns in a specific order swath_count
swath_count <- swath_count %>%
  select(-12:-15)

# Remove four columns in a specific order FtWeatherill_swath_2016
FtWeatherill_swath_2016 <- FtWeatherill_swath_2016 %>%
  select(-9:-14)

# make columns lowercase
# Get the current column names
current_names <- colnames(FtWeatherill_swath_2016)

# Convert the column names to lowercase while preserving the first letter
new_names <- sapply(current_names[1:6], function(x) {
  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
})

# Assign the new column names to the first six columns of the dataframe
colnames(FtWeatherill_swath_2016)[1:6] <- new_names

# use rbind.fill() method in R to combine data frames with different columns
swath_data <- rbind.fill(swath_binned, swath_2018, swath_count, FtWeatherill_swath_2016)

# Rename ft. wetherill to fort wetherill
swath_data$SITE[swath_data$SITE == "Ft.Weatherill"] <- "Fort Wetherill"

#still need to edit ft.wetherill dataframe

```

# Combine QUAD dataframes
```{r}

# To do: species codes transform

# use rbind.fill() method in R to combine data frames with different columns
QUAD_data <- rbind.fill(QUAD, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

# Rename ft. wetherill to fort wetherill
QUAD_data$SITE[QUAD_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"

```