### 2024-05-23 scripts/data_collation.Rmd
started script for figures

---
title: "data_collation"
author: "daniellembecker"
date: "2024-03-21"
output: html_document
---

# Use this script to clean and merge unique data types (e.g., swath, fish, UPC) from multiple years

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(lubridate)
```

# Load all data files from multiple years and seperate them into specific categories: UPC, swath, Morph, Site, QUAD, kelp, fish, and COVER
```{r}

# Set your directory path
directory <- "data/kelp_timeseries_data/"

# List files in the directory
files <- list.files(directory)

# Initialize an empty list to store data frames
data_frames <- list()

# Loop through each file
for (file in files) {
    # Read the file into a data frame
    # Adjust the read function according to your file type (e.g., read.csv, read.table)
    # Specify other parameters as needed (e.g., header = TRUE)
    file_path <- file.path(directory, file)
    df_name <- gsub("\\..*", "", file)  # Extract file name without extension
    assign(df_name, read.csv(file_path, stringsAsFactors = FALSE))
}

# Now, each file is a separate data frame with its own variable name

```

# Combine kelp dataframes
```{r}
# Kelp data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)

#delete repetative data collector columns
kelp_data <- kelp_data[ -c(11:17) ]

# Use lubridate to format date to be consistent
kelp_data$DATE <- mdy(kelp_data$DATE)

# Rename King's Beach to Kings Beach
kelp_data$SITE[kelp_data$SITE == "King's Beach"] <- "Kings Beach"

# Rename ft. weatherill to fort wetherill
kelp_data$SITE[kelp_data$SITE == "Ft Weatherill"] <- "Fort Wetherill"

```

# Combine Kelp_Morph dataframes
```{r}

# Morph data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)

# remove columns with observer and unneeded data
morph_data <- morph_data[ -c(11:14) ]

# Make all columns capitalized
colnames(morph_data) <- toupper(colnames(morph_data))

# Use lubridate to format date to be consistent
morph_data$DATE <- mdy(morph_data$DATE)

# rename moprph category columns in morph and kelp data
morph_data <- morph_data %>%
  rename(
    "BLADE_WIDTH_CM" = "WIDTH_CM")

kelp_data <- kelp_data %>%
  rename(
    "BLADE_WIDTH_CM" = "WIDTH_CM",
    "BLADE_LENGTH_CM" = "LENGTH_CM")

#MAKE STIPE LENGTH NUMERICAL INSTEAD OF CHARACTER in morph data
morph_data$STIPE_LENGTH_CM <- as.numeric(morph_data$STIPE_LENGTH_CM)

#combine kelp and morph daat
morph_data <- rbind.fill(morph_data, kelp_data)

# output updated dataframe
write.csv(morph_data, "output/cleaned_data/morph.data.clean.csv", row.names = FALSE)
```

# Kelp COVER data 
```{r}
# kelp cover dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes
# Load species names and species code data frame
species.metadata <- read.csv("data/species_code_metadata.csv")

# Remove the first column from Kelp_COVER if it's unnecessary
Kelp_COVER <- Kelp_COVER[ , -1 ]

# Function to rename columns using metadata
# Get the unique column names from the Kelp_COVER dataframe
column_names.kelp <- colnames(Kelp_COVER)

# Create a lookup table with name-to-code mappings
name_to_code.kelp <- setNames(species.metadata$code, species.metadata$name)

# Rename the columns using the lookup table
new_column_names.kelp <- sapply(column_names.kelp, function(name) {
  if (name %in% names(name_to_code.kelp)) {
    return(name_to_code.kelp[name])
  } else {
    return(name)
  }
})

# Rename the columns in Kelp_COVER
colnames(Kelp_COVER) <- new_column_names.kelp

# Melt only columns AGH:WHEL into long format
Kelp_COVER_long <- Kelp_COVER %>%
  pivot_longer(cols = AGH:CLT, names_to = "SP_CODE", values_to = "COVER")

# output updated dataframe
write.csv(Kelp_COVER_long, "output/cleaned_data/kelp.cover.data.clean.csv", row.names = FALSE)

```

# Combine fish dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes)

# remove columns with observer and unneeded data
fish_data <- fish_data[ -c(10:13) ]

# Use the actual size data to bin into the size classes to combine with other dataframe
fish$SIZE <- cut(fish$actual_size, breaks = c(0, 10, 20, 30, 40, 50, 70, 100, Inf),
               labels = c("< 10", "10_20", "20_30", "30_40", "40_50", "50_70", "70_100", "> 100"),include.lowest = TRUE, right = FALSE)

# combine updated dataframes
fish_data <- rbind.fill(fish, fish_data)

# Use lubridate to format date to be consistent
fish_data$DATE <- mdy(fish_data$DATE)

#combine with fish aggregated data
fish_data_agg <- rbind.fill(fish_data, fish_Aggregated)

# remove columns with X fromm numerical order of columns
fish_data_agg <- fish_data_agg[ -c(16) ]

# Rename ft. wetherill to fort wetherill
fish_data_agg$SITE[fish_data_agg$SITE == "Kings Beach"] <- "King's Beach"

# Rename fort weatherill
fish_data_agg$SITE[fish_data_agg$SITE == "Fort Weatherill"] <- "Fort Wetherill"

# Rename transcets to not have FW or KB before them
fish_data_agg$TRANSECT <- gsub("FW\\s*", "", fish_data_agg$TRANSECT)
fish_data_agg$TRANSECT <- gsub("KB\\s*", "", fish_data_agg$TRANSECT)

# output updated dataframe
write.csv(fish_data_agg, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)

```

# Combine UPC dataframes
```{r}

# UPC data is cleaned and ready

# use rbind.fill() method in R to combine data frames with different columns
UPC_data <- rbind.fill(UPC, UPC_2017, UPC_2018, FtWeatherill_UPC_2016, FtWeatherill_UPC_2016_CT)

# Rename ft. wetherill to fort wetherill
UPC_data$SITE[UPC_data$SITE == "Ft. Weatherill"] <- "Fort Wetherill"

# Rename fort weatherill
UPC_data$SITE[UPC_data$SITE == "Fort Weatherill"] <- "Fort Wetherill"

# remove columns with observer and unneeded data
UPC_data <- UPC_data[ -c(20:24) ]

# Make all columns capitalized
colnames(UPC_data) <- toupper(colnames(UPC_data))

# Use lubridate to format date to be consistent
UPC_data$DATE <- mdy(UPC_data$DATE)

# output updated dataframe
write.csv(UPC_data, "output/cleaned_data/UPC.data.clean.csv", row.names = FALSE)

```

# Combine site.info dataframes
```{r}
#site info cleaned metadta
# combine site collection data
site.info.dat <- kelp_siteinfo_MASTER

# output updated dataframe
write.csv(site.info.dat, "output/cleaned_data/site.info.data.clean.csv", row.names = FALSE)

```

# Combine swath dataframes
```{r}

# rename swatch category columns in swath count and swath 208 dataframe
swath_2018 <- swath_2018 %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# remove columns with observer and unneeded data
swath_2018 <- swath_2018[ -c(12:15) ]

# Make all columns capitalized
colnames(swath_2018) <- toupper(colnames(swath_2018))

#rename in and out size classes
Swath_count <- Swath_count %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# remove columns with observer and unneeded data
Swath_count <- Swath_count[ -c(12:15) ]

# Make all columns capitalized
colnames(Swath_count) <- toupper(colnames(Swath_count))

# remove columns with observer and unneeded data
FtWeatherill_Swath_2016 <- FtWeatherill_Swath_2016[ -c(9:14) ]

# Renaming values in the 'SIDE' column
FtWeatherill_Swath_2016 <- FtWeatherill_Swath_2016 %>%
  mutate(SIDE = case_when(
    SIDE == "IN 20" ~ "0-20 IN",
    SIDE == "OFF 40" ~ "40-20 OFF",
    SIDE == "IN 40" ~ "20-40 IN",
    SIDE == "OFF 20" ~ "20-0 OFF",
    TRUE ~ SIDE))  # Keep other values unchanged

# Pivoting the data from long to wide format
FtWeatherill_Swath_2016_wide <- FtWeatherill_Swath_2016 %>%
  pivot_wider(
    names_from = SIDE,
    values_from = COUNT)

# rename swatch category columns in kelp_swath 2023 and kelp swath 2022
kelp_swath_2022 <- kelp_swath_2022 %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# remove columns with observer and unneeded data
kelp_swath_2022 <- kelp_swath_2022[ -c(15:22) ]

# rename KB to King's Beach and FW to Fort Wetherill
kelp_swath_2022$SITE[kelp_swath_2022$SITE == "FW"] <- "Fort Wetherill"
kelp_swath_2022$SITE[kelp_swath_2022$SITE == "KB"] <- "King's Beach"

# rename swatch category columns in kelp_swath 2023 and kelp swath 2022
kelp_swath_2023 <- kelp_swath_2023 %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# remove columns with observer and unneeded data
kelp_swath_2023 <- kelp_swath_2023[ -c(12) ]

# rename KB to King's Beach and FW to Fort Wetherill
kelp_swath_2023$SITE[kelp_swath_2023$SITE == "FW"] <- "Fort Wetherill"
kelp_swath_2023$SITE[kelp_swath_2023$SITE == "KB"] <- "King's Beach"

# rename swatch category columns in 2019, 2020, and 2021 swath
swath_2021 <- swath_2021_DEM %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# rename KB to King's Beach and FW to Fort Wetherill
swath_2021$SITE[swath_2021$SITE == "Fort Weatherill"] <- "Fort Wetherill"
swath_2021$SITE[swath_2021$SITE == "Kings Beach"] <- "King's Beach"

# rename swatch category columns in 2019, 2020, and 2021 swath
swath_2020 <- swath_2020_DEM %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# rename Fort weatherill to Fort Wetherill
swath_2020$SITE[swath_2020$SITE == "Fort Weatherill"] <- "Fort Wetherill"

# rename swatch category columns in 2019, 2020, and 2021 swath
swath_2019 <- swath_2019_DEM %>%
  rename(
    "0-20 IN" = "X0.20.IN",
    "20-40 IN" = "X20.40.IN",
    "40-20 OFF" = "X40.20.OFF",
    "20-0 OFF" = "X20.0.OFF"
  )

# use rbind.fill() method in R to combine data frames with different columns
swath_data <- rbind.fill(swath_2018, swath_2019, swath_2020, swath_2021, Swath_count, FtWeatherill_Swath_2016_wide, kelp_swath_2022, kelp_swath_2023)

# Rename ft. wetherill to fort wetherill
swath_data$SITE[swath_data$SITE == "Ft.Weatherill"] <- "Fort Wetherill"

# Rename ft. wetherill to fort wetherill
swath_data$SITE[swath_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"

# output updated dataframe
write.csv(swath_data, "output/cleaned_data/swath.data.clean.csv", row.names = FALSE)

```

# Combine QUAD dataframes
```{r}

#make QUAD data long
# Function to rename columns using metadata
# Get the unique column names from the QUAD ataframe
column_names.quad <- colnames(QUAD)

# Create a lookup table with name-to-code mappings
name_to_code.quad <- setNames(species.metadata$code, species.metadata$name)

# Rename the columns using the lookup table
new_column_names.quad <- sapply(column_names.quad, function(name) {
  if (name %in% names(name_to_code.quad)) {
    return(name_to_code.quad[name])
  } else {
    return(name)
  }
})

# Rename the columns in Kelp_COVER
colnames(QUAD) <- new_column_names.quad

# Melt only columns AGH:WHEL into long format
QUAD_long <- QUAD %>%
  pivot_longer(cols = AGH:Oar.Weed.Total, names_to = "SP_CODE", values_to = "COUNT")

# use rbind.fill() method in R to combine data frames with different columns
QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

#remove 131-138 columns for observers not needed information
QUAD_data <- QUAD_data[ -c(14:17) ]

# Rename ft. wetherill to fort wetherill
QUAD_data$SITE[QUAD_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"

# Rename Kings Beach to King's Beach
QUAD_data$SITE[QUAD_data$SITE == "Kings Beach"] <- "King's Beach"

# output updated dataframe
write.csv(QUAD_data, "output/cleaned_data/QUAD.data.clean.csv", row.names = FALSE)

```

# Combine rugosity dataframes
```{r}
# add a YEAR column for 2021 data
# Add a new column named "year" with all values set to 2021
rugosity_2021_DEM <- data.frame(YEAR = 2021, rugosity_2021_DEM)

# use rbind.fill() method in R to combine data frames with different columns
rug_data <- rbind.fill(rugosity_2019_DEM, rugosity_2020_DEM, rugosity_2021_DEM)

#remove 22-25 columns for observers not needed information
rug_data <- rug_data[ -c(22:25) ]

# Rename Kings Beach to King's Beach
rug_data$SITE[rug_data$SITE == "Kings Beach"] <- "King's Beach"
rug_data$SITE[rug_data$SITE == "KB"] <- "King's Beach"

# rename rugosity data columns
rug_data <- rug_data %>%
  rename(
    "0M" = "X0M",
    "4M" = "X4M",
    "8M" = "X8M",
    "12M" = "X12M",
    "16M" = "X16M",
    "20M" = "X20M",
    "24M" = "X24M",
    "28M" = "X28M",
    "32M" = "X32M",
    "36M" = "X36M",
    "40M" = "X40M", 
    "DEPTH.0M" = "DEPTH...0M..ft.",
    "DEPTH.40M" = "DEPTH....40M.",
    "VIS.M" = "VIS..M."
    )

# output updated dataframe
write.csv(rug_data, "output/cleaned_data/rugosity.data.clean.csv", row.names = FALSE)

```