### 2024-06-03 scripts/heatwave_script.Rmd
made figure 3

---
title: "RI_heatwave"
author: "daniellembecker"
date: "1/15/2022"
output:
  pdf_document: default
  html_document: default
adapted from: amandachiachi
---

#OISST_data
#Adapted by Amanda Chiachi
#11/13/2019

#### Setup ##########

#load libraries
```{r}
library(dplyr)
library(ggplot2)
library(ncdf4)
library(heatwaveR)# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(lubridate)
library(scales)
library(maps)

```

# The information for the NOAA OISST data
```{r}
rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")
```

# This function expects the user to provide it with a start and end date
# It then downloads and prepares the data
# we are downloading only the SST data out of several variables 
# spatial extent of latitude for ft wetheril
```{r}
# Define the OISST_sub_dl function
OISST_sub_dl <- function(time_df){
  print(time_df)  # Debug print to check the structure of time_df

  # Convert start and end dates to character strings
  start_date <- as.character(time_df$start)
  end_date <- as.character(time_df$end)

  print(paste("Start date:", start_date))
  print(paste("End date:", end_date))

  OISST_dat <- griddap(datasetx = "ncdcOisst21Agg_LonPM180", 
                       url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                       time = c(start_date, end_date), 
                       zlev = c(0, 0),
                       latitude = c(41.47944444, 41.47944444),
                       longitude = c(-71.35611111, -71.35555556),
                       fields = "sst")$data %>% 
    mutate(time = as.Date(stringr::str_remove(time, "T00:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst) %>% 
    select(lon, lat, t, temp) %>% 
    na.omit()

  return(OISST_dat)
}

```

#server doesnt like more than 9 years of consecutive data 
#creates a data frame to download multiple batches of the data 
# allows us to automate the entire download 
# Date download range by start and end dates per year
```{r}
# Define the date ranges for data download
ft_weth_years <- data.frame(date_index = 1,
                            start = as.Date("2016-01-01"),
                            end = as.Date("2023-12-29"))

# Execute the data download and processing
system.time(
  OISST_data <- ft_weth_years %>% 
    group_by(date_index) %>% 
    group_modify(~OISST_sub_dl(.x)) %>% 
    ungroup() %>% 
    select(lon, lat, t, temp)
)

#format date and time, lubridate package format
OISST_data$t <- ymd(OISST_data$t)

#subset temp data if needed

heatwave.temps.OISST <- OISST_data %>%
  filter(lubridate::month(t) %in% c(3:4))

heatwave.temps_filter.OISST <- heatwave.temps.OISST %>%
  filter(lubridate::year(t) %in% c(2016:2024))

write.csv(heatwave.temps_filter.OISST, "output/ft.wetherill.heatwave.temps.OISST.csv")

```

# Detect the events in a time series
# "...must be at least three years to calculate thresholds"
# create your time series data 
```{r}
ts.OISST <- ts2clm(OISST_data, climatologyPeriod = c("2002-01-01", "2021-11-01"))
mhw.OISST <- detect_event(ts.OISST)
```

# View just a few metrics
# The function will return a line plot indicating the climatology, threshold and temperature, with the hot or cold events that meet the specifications of Hobday et al. (2016) shaded in as appropriate.
# An anomalously warm event is considered to be a MHW if it lasts for five or more days, with temperatures warmer than the 90th percentile based on a 30-year historical baseline period.
```{r}
#make data frame with marine heatwave events over time in Mo'orea

mhw.record.OISST <- mhw.OISST$event %>% 
  dplyr::ungroup() %>%
  dplyr::select(event_no, duration, date_start, date_end, date_peak, intensity_max, intensity_cumulative, intensity_mean, rate_onset, rate_decline) %>% 
  dplyr::arrange(-intensity_max); mhw.record.OISST

write.csv(mhw.record.OISST, "RAnalysis/output/heatwave/marine.heatwave.record.OISST.csv")

#format date and time
mhw.OISST$Date<- mdy(mhw.OISST$Date)

#create a line plot of marine heatwave data, can change date range for more specific times
#2019 heatwave scenarios
event_line(mhw.OISST, spread = 50, metric = "intensity_max", 
           start_date = "2019-03-01", end_date = "2019-05-01", x_axis_title = "Date") +
    scale_x_date(date_breaks = "1 day", labels = date_format("%b-%d")) + 
    theme_classic()+
    theme(axis.text.x = element_text(angle = 90))

```